---
title: 'Assessment Part 1: Dates, Times, and Text Mining'
author: "Werner Dassuncao"
date: "```{r} format(Sys.Date(), '%d %B, %Y.')`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assessment Part 1: Dates, Times, and Text Mining
### Solutions by Werner Dassuncao

This assessment reviews several concepts about dates, times, and text mining. In part 1 on this page, you will practice extracting and manipulating dates in real datasets. In part 2 on the next page, you will walk through a sentiment analysis of a novel using steps covered in the previous section.

Use the following libraries and options for coding questions:

```{r}
library(dslabs)
library(tidyverse)
library(lubridate)
options(digits = 3)    # 3 significant digits
```

IMPORTANT: Some of these exercises use dslabs datasets that were added in a July 2019 update. Make sure your package is up to date with the command install.packages("dslabs").

### Question 1
Which of the following is the standard ISO 8601 format for dates?


MM-DD-YY
YYYY-MM-DD  ***** Correct answer: This is proper ISO 8601 formatting for dates.
YYYYMMDD
YY-MM-DD

### Question 2

Which of the following commands could convert this string into the correct date format?

```{r}      
dates <- c("09-01-02", "01-12-07", "02-03-04")

ymd(dates) 
        
mdy(dates)
        
dmy(dates)
```

Answer:
It is impossible to know which format is correct without additional information.

Correct:The formatting of these dates is ambiguous. They could be formatted as ymd, mdy or dmy. We need more information about our data to be able to select the correct command.


### Question 3

Load the brexit_polls data frame from dslabs:

`r data(brexit_polls)`

  
How many polls had a start date (startdate) in April (month number 4)?
```{r}
# Using the library lubridate and function as.Date:
library(lubridate)
p <- brexit_polls %>% filter(startdate >= as.Date('2016-04-01') & startdate <= as.Date('2016-04-30'))
head(p)
nrow(p)

#Another option:
sum(month(brexit_polls$startdate) == 4)


```
 
Use the round_date() function on the enddate column with the argument unit="week". How many polls ended the week of 2016-06-12?
Read the documentation to learn more about round_date().

```{r}
sum(round_date(brexit_polls$enddate, unit = "week") == "2016-06-12")
```

### Question 4

Use the weekdays() function from lubridate to determine the weekday on which each poll ended (enddate).

On which weekday did the greatest number of polls end?

```{r}
brexit_polls %>%  
  mutate(weekday = weekdays(enddate)) %>%
  count(weekday) %>%
  arrange(desc(weekday))
  

#alternative answer:
table(weekdays(brexit_polls$enddate))
  
```



### Question 5

Load the movielens data frame from dslabs.

    
`r data(movielens)`

  
This data frame contains a set of about 100,000 movie reviews. The timestamp column contains the review date as the number of seconds since 1970-01-01 (epoch time).

Convert the timestamp column to dates using the lubridate as_datetime() function.

Which year had the most movie reviews?

```{r}
str(movielens)

movielens %>% 
  mutate(d_year = year(as_datetime(timestamp))) %>%
  group_by(d_year) %>%
  count(d_year) %>%
  arrange(desc(n))



#Alternative solution:
dates <- as_datetime(movielens$timestamp) #grabbing the dates
reviews_by_year <- table(year(dates))    # count reviews by year
names(which.max(reviews_by_year))    # name of year with most reviews

```
 
Which hour of the day had the most movie reviews?
```{r}

movielens %>% 
  mutate(review_hour = hour(as_datetime(timestamp, tz = "UTC"))) %>%
  group_by(review_hour) %>%
  count(review_hour) %>%
  arrange(desc(n))


# Alternative solution:
reviews_by_hour <- table(hour(dates))    # count reviews by hour
names(which.max(reviews_by_hour))    # name of hour with most reviews

```











