---
title: "Assessment Part 2: String Processing Part 3"
author: "Werner Dassuncao"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import raw Brexit referendum polling data from Wikipedia:

```{r}
library(rvest)
library(tidyverse)
library(stringr)
url <- "https://en.wikipedia.org/w/index.php?title=Opinion_polling_for_the_United_Kingdom_European_Union_membership_referendum&oldid=896735054"
tab <- read_html(url) %>% 
  html_nodes("table")

polls <- tab[[5]] %>%
  html_table(fill = TRUE)
```

&nbsp;
You will use a variety of string processing techniques learned in this section to reformat these data.

&nbsp;

## Question 5

Some rows in this table do not contain polls. You can identify these by the lack of the percent sign (%) in the Remain column.

Update polls by changing the column names to c("dates", "remain", "leave", "undecided", "lead", "samplesize", "pollster", "poll_type", "notes") and only keeping rows that have a percent sign (%) in the remain column.

```{r}
polls <- polls %>%
  filter(str_detect(Remain, "%"))

colnames(polls) <-  c('dates', 'remain', 'leave', 'undecided', 'lead', 'samplesize', 'pollster', 'poll_type', 'notes')
head(polls)
```


```{r Alternative code, echo=FALSE}
#names(polls) <- c("dates", "remain", "leave", "undecided", "lead", "samplesize", "pollster", "poll_type", "notes")
#polls <- polls[str_detect(polls$remain, "%"), -9]
#nrow(polls)
```



How many rows remain in the polls data frame?
```{r}
nrow(polls)
```

## Question 6

The remain and leave columns are both given in the format "48.1%": percentages out of 100% with a percent symbol.

Which of these commands converts the remain vector to a proportion between 0 and 1?
Check all correct answers.

```{r}
as.numeric(str_remove(polls$remain, "%"))

as.numeric(polls$remain)/100

parse_number(polls$remain)

#non-numeric argument to binary operator: commented out
#str_remove(polls$remain, "%")/100

#Correct
as.numeric(str_replace(polls$remain, "%", ""))/100


#Correct
parse_number(polls$remain)/100
```

&nbsp;

## Question 7

The undecided column has some "N/A" values. These "N/A"s are only present when the remain and leave columns total 100%, so they should actually be zeros.

Use a function from stringr to convert "N/A" in the undecided column to 0. The format of your command should be function_name(polls$undecided, "arg1", "arg2").

What function replaces function_name?
  
```{r}
polls <- polls %>%
  mutate(undecided = str_replace_all(undecided, "N/A", "0"))

polls
```
  
What argument replaces arg1?
Omit the quotation marks.

Answer: "N/A"

What argument replaces arg2?
Omit the quotation marks.

Answer: 0

&nbsp;
&nbsp;


## Question 8

The dates column contains the range of dates over which the poll was conducted. The format is "8-10 Jan" where the poll had a start date of 2016-01-08 and end date of 2016-01-10. Some polls go across month boundaries (16 May-12 June).

The end date of the poll will always be one or two digits, followed by a space, followed by the month as one or more letters (either capital or lowercase). In these data, all month abbreviations or names have 3, 4 or 5 letters.

Write a regular expression to extract the end day and month from dates. Insert it into the skeleton code below:

```{r}      
#temp <- str_extract_all(polls$dates, ____________)
#end_date <- sapply(temp, function(x) x[length(x)]) # take last element (handles polls that cross month boundaries)
#end_date
```
    
Which of the following regular expressions correctly extracts the end day and month when inserted into the blank in the code above?
Check all correct answers.

```{r}
#"\\d?\\s[a-zA-Z]?"
pattern4 <- "\\d+\\s[a-zA-Z]+"
#"\\d+\\s[A-Z]+"
pattern3 <- "[0-9]+\\s[a-zA-Z]+"
pattern2 <- "\\d{1,2}\\s[a-zA-Z]+"
#"\\d{1,2}[a-zA-Z]+"
pattern1 <- "\\d+\\s[a-zA-Z]{3,5}"
```


```{r}      
temp <- str_extract_all(polls$dates, pattern1)
end_date <- sapply(temp, function(x) x[length(x)]) # take last element (handles polls that cross month boundaries)
end_date
```
Final note: All patterns 1-4 are usable for grabbing the end day and month









