---
title: "Section 4: Dates, Times and Text Mining"
author: "Werner Dassuncao" 
date: "`r format(Sys.Date(), '%d %B, %Y.')`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Key points
&nbsp;
Dates are a separate data type in R.The tidyverse includes functionality for dealing with dates through the lubridate package. 

Extract the year, month and day from a date object with the year(), month() and day() functions.

Parsers convert strings into dates with the standard **YYYY-MM-DD** format (ISO 8601 format). Use the parser with the name corresponding to the string format of year, month and day (ymd(), ydm(), myd(), mdy(), dmy(), dym()).

Get the current time with the Sys.time() function. Use the now() function instead to specify a time zone.

You can extract values from time objects with the hour(), minute() and second() functions.

Parsers convert strings into times (for example, hms()). Parsers can also create combined date-time objects (for example, mdy_hms()).
&nbsp;

```{r echo=TRUE}

# inspect the startdate column of 2016 polls data, a Date type
library(tidyr)
library(tidyverse)
library(dslabs)
data("polls_us_election_2016")
polls_us_election_2016$startdate %>% head
class(polls_us_election_2016$startdate)

# Unix epoch is 00:00:00 UTC on 1 January 1970, dates are represented as days since epoch:
as.numeric(polls_us_election_2016$startdate) %>% head

# ggplot is aware of dates
polls_us_election_2016 %>% filter(pollster == "Ipsos" & state =="U.S.") %>%
  ggplot(aes(startdate, rawpoll_trump)) +
  geom_line()

# lubridate: the tidyverse date package
library(lubridate)

# select some random dates from polls
set.seed(2)
dates <- sample(polls_us_election_2016$startdate, 10) %>% sort
dates

# extract month, day, year from date strings
data.frame(date = dates, 
           month = month(dates),
           day = day(dates),
           year = year(dates))

month(dates, label = TRUE)    # extract month label

# ymd works on mixed date styles
x <- c(20090101, "2009-01-02", "2009 01 03", "2009-1-4",
       "2009-1, 5", "Created on 2009 1 6", "200901 !!! 07")
ymd(x)

# different parsers extract year, month and day in different orders
x <- "09/01/02"
ymd(x)
mdy(x)
ydm(x)
myd(x)
dmy(x)
dym(x)

now()    # current time in your time zone
now("GMT")    # current time in GMT
now() %>% hour()    # current hour
now() %>% minute()    # current minute
now() %>% second()    # current second

# parse time
x <- c("12:34:56")
hms(x)

#parse datetime
x <- "Nov/2/2012 12:34:56"
mdy_hms(x)
```

















